### 2025-05-20 tex/FSCD25/rebuttal.txt
Aftermath of the FSCD rejection

We thank the reviewers for their detailed and constructive feedback.

Reviewer 1:

>Lack of focus on metaproperties about object type theory

The interpretation into the standard model is described in Section 4.2, albeit
very briefly due to a lack of space. But it demonstrates that our approach can
remove noise in definitions and proofs, with the details available in the Agda
formalisation, as stated in Section 4.2.

>Practical effectiveness of the approach

We didn't manage to remove all transports in the case of inductive types, as we
acknowledge in Section 3.5. However, we still improve on the number of
transports needed for types and introduction forms.  We concentrate on
eliminating transports along type substitutions, which introduce overhead early
in the formalisation. Addressing this specific source of transports hopefully
serves as a use case of QIIRTs and a step towards a solution to the transport
hell. Our preliminary substitution calculus NbE experiment that the reviewer
mentions indicate that the QIIRT-based definition reduces the code size by
approximately 20% (from 229 to 182 LoC), for this simple case.

>It seems like [your GitHub repository] contains more, such as work-in-progress
>proof of NbE. Is there a reason to exclude that from your future work?

The current experiments are small in terms of code size, and thus we didn't
think it was worthwhile to mention them in the paper. But we of course hope to
turn them into larger projects at a later stage. Since the QIIRT definition is
(syntactically) isomorphic to the QIIT one, and the eliminator only introduces
additional derived equations, we expect existing meta-theoretical results based
on QIITs to transfer with minimal changes.

Reviewer 2:

>Do you have any (even vague) idea of how your generalisation of IRT would fare
>with any of the justification for induction-recursion?

It seems plausible that Dybjer and Setzer's set-theoretic model ([20] in our
paper) could be adapted to justify this generalisation. If so, it should also
be straightforward to extend this model to also include equality constructors,
since quotients are easy in set theory.

>Can you explain how you derive your notion of eliminator?
>In particular, what (if any) is the difference between that eliminator and
>that for the standard QIIT?
>And do you have a justification/intuition for why you require that your model
>satisfies only propositionally the rules which are definitional in the QIIRT?

The eliminator for a QIIRT is derived following the general QIIT schema, where
each function clause in the recursive part of the QIIRT definition corresponds
to an equality constructor.

Since we implement some constructors as recursive functions rather than
constructors, the associated computation rules become propositional equalities
in the eliminator (Proposition 4.2). These propositional equalities cannot be
stated as definitional equalities, since there is no way to talk about
definitional equality within type theory.

For the same reason, models cannot be required to satisfy definitional equalities for
these clauses.

>Can you explain why you are not able to have a notion of term-level
>substitution defined solely as a recursive function?
>Why you chose to have composition and other operation not be themselves
>defined? And why substitution is defined by induction on the structure of
>substitution, rather than on that of terms?

We do not define term-level substitution recursively, as doing so does not seem
to reduce the number of transports needed, beyond the "smart constructor"
[–]Tm– (Section 3.3). Defining more operations recursively would increase the
number of proof obligations such as verifying confluence.

The smart constructor is defined by induction on the structure of substitution,
because the clauses based on the substitution structure are exactly those needed
to align with the definitional behaviour of type substitution.

>Can you comment on the relation with research on explicit substitution
>calculi? In particular, do you view any fundamental obstruction to/reason not
>to adapt the complete rewriting system of Schäfer, Smolka and Tebbi, which
>seems to solve your issue in section 3.5, to a typed setting?

The work of Schäfer et al. employs a normalisation function into normal
expressions (see Section 6 and Theorem 32, op. cit.). In our setting, this
would correspond to replacing equality constructors with a normalisation
function, similar to the QIRT definition of join lists in Section 2.2 and the
work by Bense et al. Adapting their approach to a type theory with universes
and inductive types seems useful but likely requires a definition that departs
far from the typical CwF-based definition of type theory.

Reviewer 3:

>In your formalisation there is less transport hell, but is there any kind of
>new hell? E.g. "implicit argument hell" [...].

Indeed, some implicit arguments must be given explicitly in the QIIRT version.
Compared to the work saved from dealing with transports explicitly, we believe
it is a good deal, since inserting implicit arguments does not complicate the
equational theory.